Geometric Filters for Metric Spaces

### 2018-11-27 paper/draft.tex
interim

We tested the algorithm for $\eps \in \Set{0.01, 0.1, 0.2, 0.5}$ on the following sets of points in dimensions $d = 2,3,4,5$:
\begin{enumerate}
    \item In the \textbf{uniform} test set points are sampled uniformly at
        random from the unit cube in $\mathbb{R}^d$.
    \item In the \textbf{normal} test set points are sampled from the standard
        normal distribution in $\mathbb{R}^d$.
    \item In the \textbf{clustered} test set we first sample cluster centers uniformly 
        at random from $[0,10000]^d$, and then we add normally distributed noise around
        each of the centers. The number of clusters is chosen so that each cluster
        contains 50 points.
    \item The test set \textbf{exp} consists of points of the form $(2^{\xi_1}, \dots, 2^{xi_d})$,
        where $\xi_i$'s are i.i.d. random variables with uniform distribution on $[1,25]$.
\end{enumerate}
In all experiments the algorithms that we tested compared in the same way,
so we only present results for the \textbf{uniform} point set in dimension 2.
%Let $e$ denote the number of edges of a spanner on $n$ vertices;
%in the plots we visualize the ratios $e / { n \choose 2}$ and $e / n$.
%We refer to the first ratio as \textit{density} and to the second one
%as \textit{growth}. Note that 
The results are summarized in the plots \ref{plot1}, \ref{fig:spanner_plot2}.
The first plot shows the number of edges of the spanner.
We can see that, even though none of the blind spanners can produce
spanners of the same quality (i.e., sparseness) as the standard greedy algorithm,
\textsc{BlindGreedy} and all variants of \textsc{BlindRandom}
perform significantly better than both variants of \textsc{BlindQuasiSorted}.
it is important to remember that for all blind spanners the number of edges
is equal to the number of distances that were actually computed, while
for non-blind greedy spanner the latter number is always $n \choose 2$.
The second plot shows the ratio of the number of edges to the number of points.
The ideal behavior is demonstrated by the non-blind greedy spanner,
for which this ratio stays practically constant, confirming the linear growth.
None of the blind algorithms seems to have this property, but among them
the blind greedy spanner is the best one. If we assume that the number of edges
is proportional to $n^\alpha$, then we can try to estimate $\alpha$ by 
linear regression.  We give in the table \ref{tbl:regr_coeff_spanner}
the maximal exponents $\alpha$ for each of the algorithms,
and it is worth mentioning that even for the quasi-sorted shaker variant 
the exponent seems to be bounded away from 2.

>>>>

\begin{figure}[ht]
    \label{fig:spanner-sparseness}
    \includegraphics[width=0.7\textwidth]{edges_to_points_ratio_dim_2_normal_points.png}
    \caption{Number of edges in blind spanners generated by different variants of the blind
    algorithm. Greedy non-blind algorithm is included for comparison. The plot is for normally distributed points
    in dimension 2, $\eps = 0.1$.}
\end{figure}

\begin{figure}[ht]
    \label{fig:blind-greedy-only}
    \includegraphics[width=0.7\textwidth]{blind-greedy-dims.png}
    \caption{}
\end{figure}

We tested the algorithm for $\eps \in \Set{0.01, 0.1, 0.2, 0.5}$ on the following sets of points in dimensions $d = 2,3,4,5$:
\begin{enumerate}
    \item In the \textbf{uniform} test set points are sampled uniformly at
        random from the unit cube in $\mathbb{R}^d$.
    \item In the \textbf{normal} test set points are sampled from the standard
        normal distribution in $\mathbb{R}^d$.
    \item In the \textbf{clustered} test set we first sample cluster centers uniformly 
        at random from $[0,10000]^d$, and then we add normally distributed noise around
        each of the centers. The number of clusters is chosen so that each cluster
        contains 50 points.
    \item The test set \textbf{exp} consists of points of the form $(2^{\xi_1}, \dots, 2^{xi_d})$,
        where $\xi_i$'s are i.i.d. random variables with uniform distribution on $[1,25]$.
\end{enumerate}
In all experiments the algorithms that we tested compared in the same way,
so we only present results for the \textbf{uniform} point set in dimension 2.
%Let $e$ denote the number of edges of a spanner on $n$ vertices;
%in the plots we visualize the ratios $e / { n \choose 2}$ and $e / n$.
%We refer to the first ratio as \textit{density} and to the second one
%as \textit{growth}. Note that

The results are summarized in the plot \ref{fig:spanner-sparseness}.
The first plot shows the number of edges of the spanner.
We can see that, even though none of the blind spanners can produce
spanners of the same quality (i.e., sparseness) as the standard greedy algorithm,
\textsc{BlindGreedy} and all variants of \textsc{BlindRandom}
perform significantly better than both variants of \textsc{BlindQuasiSorted}.
it is important to remember that for all blind spanners the number of edges
is equal to the number of distances that were actually computed, while
for non-blind greedy spanner the latter number is always $n \choose 2$.
The second plot shows the ratio of the number of edges to the number of points.
The ideal behavior is demonstrated by the non-blind greedy spanner,
for which this ratio stays practically constant, confirming the linear growth.
None of the blind algorithms seems to have this property, but among them
the blind greedy spanner is the best one. If we assume that the number of edges
is proportional to $n^\alpha$, then we can try to estimate $\alpha$ by 
linear regression.  We give in the table \ref{tbl:regr_coeff_spanner}
the maximal exponents $\alpha$ for each of the algorithms,
and it is worth mentioning that even for the quasi-sorted shaker variant 
the exponent seems to be bounded away from 2.

As for different variants of the \textsc{BlindRandomBadRatio} algorithm,
we note that their performance is almost the same, and the algorithm works
significantly better than \textsc{QuasiSorted} variants, but obviously
worse than the blind greedy variant. There is a consistent, though small,
difference between the variants that do not force lower bounds first and
the other two variants of the BlindRandomBadRatio.

We also look at higher dimensions, but the situation there is more
disappointing. We show the results for the best algorithm, \textsc{BlindGreedy},
in dimensions 2, 3 and 4 in the plot \ref{fig:blind-greedy-only}.
We can see that already in dimension 4 the best available algorithm
for the uniformly distributed point sets the blind greedy variant
produces almost a complete graph up to 700 points.
